---
title: "string practice"
author: "ims"
format: html
---

Task 3: Practice with strings

Create a new .qmd in your R folder called string_practice.qmd

Copy all contents of the html table below to your clipboard:
date	building	alarm_message
2020-03-14	Engineering-North	10:02am -- HVAC system down, facilities management alerted
2020-03-15	Bren Hall	8:24am -- Elevator North out of service
2020-04-10	Engineering-South	12:41am -- Fire alarm, UCSB fire responded and cleared
2020-04-18	Engr-North	9:58pm -- Campus point emergency siren, UCPD responded
Back in your string_practice.Rmd, create a new code chunk


With your cursor in your code chunk, go up to Addins in the top bar of RStudio. From the drop-down menu, choose ‘Paste as data frame’. Make sure to add code to store the data frame as alarm_report
```{r}

alarm_report <- data.frame(
  stringsAsFactors = FALSE,
              date = c("2020-03-14", "2020-03-15", "2020-04-10", "2020-04-18"),
          building = c("Engineering-North",
                       "Bren Hall","Engineering-South","Engr-North"),
     alarm_message = c("10:02am -- HVAC system down, facilities management alerted",
                       "8:24am -- Elevator North out of service",
                       "12:41am -- Fire alarm, UCSB fire responded and cleared",
                       "9:58pm -- Campus point emergency siren, UCPD responded")
)
```

Practice working with strings by writing code to update alarm_report as follows (these can be separate, or all as part of a piped sequence):

Replace the “Engr” with “Engineering” in the building column
Separate the building column into two separate columns, building and wing, separated at the dash
Only keep observations with the word “responded” in the alarm_message column
Separate the message time from the rest of the message by separating at --
Convert the date column to a Date class using lubridate

```{r}
update_alarm_report <- alarm_report |> 
  mutate(building = str_replace(building, pattern = "Engr", replacement = "Engineering")) |> 
  separate(building, c("building", "wing")) |> 
  filter(str_detect(alarm_message, "responded")) |> 
  separate(alarm_message, c("time", "message"), sep = "--") |> 
  mutate(date = lubridate::date(date))  
  
update_alarm_report
```

